


## Background


## Mean Total Number of Steps taken per day

```{r}
activitydata <- read.csv("activity.csv", header=TRUE, sep=",")
totalsteps <- sum(activitydata$steps,na.rm=TRUE)
ndays <- length(unique(activitydata$date))
```

Everyday an average of `r round(totalsteps/ndays)` steps were taken.


# Distribution of the total number of steps taken everyday.


```{r}
stepactivity <- activitydata[,c(1,2)]
filterstepactivity <- na.omit(stepactivity)

histagg <- aggregate(filterstepactivity$steps, by=list(date=filterstepactivity$date), FUN=sum)

xt <- as.POSIXct(strptime(histagg$date, "%Y-%m-%d"))
yt <- histagg$x
barplot(height = yt , names.arg= xt,  ylab = "Steps", xlab="Dates",col = "red")
```

# Mean and median of the steps taken everyday.

```{r}
summary(histagg)
```


## What is the average daily activity pattern?

```{r}
stepintervalactivity <- na.omit(activitydata[,c(1,3)])
timeagg <- aggregate(x=stepintervalactivity$steps, by=list(interval=stepintervalactivity$interval), FUN=mean)


#> stepintervalactivity <- na.omit(activitydata[,c(1,3)])
#> timeagg <- aggregate(stepintervalactivity$steps, by=list(interval=stepintervalactivity$interval), FUN=mean)

convert_time = function(x) {

y <- x %% 100
q <-  x %/% 100

return ( y + 60 *q )
}

xt <- convert_time(timeagg$interval)
plot(timeagg$x ~ xt, type = "l", ylab = "Average no of steps", xlab = "Time of the day ( no of 5 minute intervals)")
```

```{r}
head(timeagg[order(timeagg$x, decreasing = TRUE),])
```

The time interval around `r convert_time(835)`  minutes (  8:35 AM ) has the highest number of steps.

##  Imputing missing values

```{r}
missingValues <-  nrow(activitydata) - nrow(na.omit(activitydata))
```

There are `r missingValues` missing values in the dataset.

```{r}
timeagg <- aggregate(stepintervalactivity$steps, by=list(interval=stepintervalactivity$interval), FUN=mean)

replace_null = function(a,b ) {
if ( is.na(a) ) return ( b )
return ( a)
}

activity2 <- merge(activitydata, timeagg, by.x = "interval", by.y ="interval")



#subvvxtsagg <- apply.daily(subvvxts, FUN=mean)
#subvvxts <- xts(subvv[,-1], order.by=as.POSIXct(subvv$date))


replace_null_with_mean = function(x, y) {
     len <- nrow(x)
	df = data.frame(steps = numeric(len))
     for ( i in 1:len){
	df$steps[i] <- round(replace_null( x[i,2], x[i,4]))
     }
     return ( df)
}


nn <- replace_null_with_mean(activity2, timeagg)

vv <- data.frame(activity2[,c(1,3)], nn)

vvdateagg <- aggregate(vv$steps, by=list(date=vv$date), FUN=sum)



xtc <- as.POSIXct(strptime(vvdateagg$date, "%Y-%m-%d"))
ytc <- vvdateagg$x
barplot(height = ytc , names.arg= xtc,  ylab = "Steps", xlab="Dates",col = "red")

summary(vvdateagg)

ytdiff <- ytc - yt

barplot(height = ytdiff , names.arg= xtc,  ylab = "Steps", xlab="Dates",col = "red")
```

## Difference in activity Patterns between weekdays and weekends.
```{r}
wd <- weekdays(as.Date(vvdateagg$date))
wdaydata <- data.frame(vvdateagg, wd)
weekdayagg <- aggregate(wdaydata$x, by=list(day=wdaydata$wd), FUN=mean)

xtw <- weekdayagg$day
ytw <- weekdayagg$x

barplot(height = ytw , names.arg= xtw,  ylab = "Steps", xlab="Dates",col = "red")
weekdayagg 
```


There is slightly more activity on weekends as compared to weekdays.
 
