


## Background

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The 
data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the 
number of steps taken in 5 minute intervals each day.
The data has been analyzed to answer the following questions.

 +  What is the mean of the total number of steps taken each day.
 +  What is the average daily activity pattern
 +   What is the impact of imputing missing values in the data
 +  What is the difference in activity comparing weekdays and weekends.


## Mean Total Number of Steps taken per day

```{r echo=FALSE}
activitydata <- read.csv("activity.csv", header=TRUE, sep=",")
totalsteps <- sum(activitydata$steps,na.rm=TRUE)
ndays <- length(unique(activitydata$date))
```

Everyday an average of `r round(totalsteps/ndays)` steps were taken.


# Distribution of the total number of steps taken everyday.


```{r echo=FALSE}
stepactivity <- activitydata[,c(1,2)]
filterstepactivity <- na.omit(stepactivity)

histagg <- aggregate(filterstepactivity$steps, by=list(date=filterstepactivity$date), FUN=sum)

xt <- as.POSIXct(strptime(histagg$date, "%Y-%m-%d"))
yt <- histagg$x
barplot(height = yt , names.arg= xt, main="Steps taken on any day",  ylab = "Steps", xlab="Dates",col = "red")
```

# Mean and median of the steps taken everyday.

```{r echo=FALSE}
summary(histagg)
```


## What is the average daily activity pattern?

```{r echo=FALSE}
stepintervalactivity <- na.omit(activitydata[,c(1,3)])
timeagg <- aggregate(x=stepintervalactivity$steps, by=list(interval=stepintervalactivity$interval), FUN=mean)


#> stepintervalactivity <- na.omit(activitydata[,c(1,3)])
#> timeagg <- aggregate(stepintervalactivity$steps, by=list(interval=stepintervalactivity$interval), FUN=mean)

convert_time = function(x) {

y <- x %% 100
q <-  x %/% 100

return ( y + 60 *q )
}

xt <- convert_time(timeagg$interval)
plot(timeagg$x ~ xt, type = "l", main = "Average number of steps taken over a 5 minute interval in a day",ylab = "Average no of steps", xlab = "Time of the day ( no of 5 minute intervals)")
```

```{r echo=FALSE}
head(timeagg[order(timeagg$x, decreasing = TRUE),])
```

The time interval around `r convert_time(835)`  minutes (  8:35 AM ) has the highest number of steps.

##  Imputing missing values

```{r echo = FALSE}
missingValues <-  nrow(activitydata) - nrow(na.omit(activitydata))
```

There are `r missingValues` missing values in the dataset.

```{r echo=FALSE}
timeagg <- aggregate(stepintervalactivity$steps, by=list(interval=stepintervalactivity$interval), FUN=mean)

replace_null = function(a,b ) {
if ( is.na(a) ) return ( b )
return ( a)
}

activity2 <- merge(activitydata, timeagg, by.x = "interval", by.y ="interval")



#subvvxtsagg <- apply.daily(subvvxts, FUN=mean)
#subvvxts <- xts(subvv[,-1], order.by=as.POSIXct(subvv$date))


replace_null_with_mean = function(x, y) {
     len <- nrow(x)
	df = data.frame(steps = numeric(len))
     for ( i in 1:len){
	df$steps[i] <- round(replace_null( x[i,2], x[i,4]))
     }
     return ( df)
}


nn <- replace_null_with_mean(activity2, timeagg)

vv <- data.frame(activity2[,c(1,3)], nn)

vvdateagg <- aggregate(vv$steps, by=list(date=vv$date), FUN=sum)



xtc <- as.POSIXct(strptime(vvdateagg$date, "%Y-%m-%d"))
ytc <- vvdateagg$x
barplot(height = ytc , names.arg= xtc,  main = "Average daily number of steps with inputed null values",ylab = "Steps", xlab="Dates",col = "red")

summary(vvdateagg)

ytdiff <- ytc - yt

barplot(height = ytdiff , names.arg= xtc,  main = "Impact of imputed Values on the Number of steps per day",ylab = "Steps", xlab="Dates",col = "red")
```

## Difference in activity Patterns between weekdays and weekends.
```{r echo=FALSE}
wd <- weekdays(as.Date(vvdateagg$date))
wdaydata <- data.frame(vvdateagg, wd)
weekdayagg <- aggregate(wdaydata$x, by=list(day=wdaydata$wd), FUN=mean)

xtw <- weekdayagg$day
ytw <- weekdayagg$x

barplot(height = ytw , names.arg= xtw, main= "Average number of steps taken on any day of the week", ylab = "Steps", xlab="Dates",col = "red")
weekdayagg
mean_wday <- (sum(ytw) - ytw[3] - ytw[4])/5 
mean_wend <- (ytw[3] + ytw[4])/2
```


 There is slightly more activity on weekends as compared to weekdays.
 `r round(mean_wend - mean_wday)` additional steps were taken  on weekends.
 
